Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies

### 2021-12-20 ms/sections/discussion.tex
done with revision

Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers)\textcolor{red}{. Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis), and (3) \textit{A. sagrei} is highly mobile \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.} \textcolor{olive}{, and we would have expected gene flow to cause a more homogeneous distribution of color phenotypes within islands. While little is known about the cruising range of individuals from our study populations (but see \citealt{Steinberg2017, Kamath2018} for other systems), \textit{A. sagrei} are polygynandrous (both males and females mate with multiple mates, \citealt{Kamath2017, Kamath2018a, Kamath2018}), thus offering opportunity for gene flow, especially given that lizards were distributed continuously and at high densities within the islands we sampled. Consistent with that, while populations from different islands were monophyletic, individuals within islands were not monophyletic with respect to habitat based on mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis).}\\

@@

\textcolor{red}{We found no evidence for a role of genetic drift in explaining the observed patterns either. First, \textit{A. sagrei} was distributed across the islands continuously, usually at relatively high population densities, rather than in small and isolated populations where drift might be expected to have a strong effect. Second, we found no evidence of isolation-by-distance except on Abaco, which was sampled at the largest geographical scale, with sites nearly a hundred kilometers apart from each other. Hence, while isolation-by-distance may explain long-range differences on this island, most of the differences among habitats across the rest of the sampling region are unlikely to be explained by genetic drift, as habitats were often in close proximity (less than 10km).}\\

@@

In this study, we found larger differences among than within islands, a pattern already reported and linked to climatic conditions \citep{Driessens2017} and to densities of predators and of anole congeners \citep{Vanhooydonck2009, Baeckens2018}. Differences among habitats within islands, however, are still difficult to account for. Remaining hypotheses may include, for example, runaway sexual selection \textcolor{olive}{(i.e. arbitrary preferences of females for some colors over others, \citealt{Andersson1994})} \textcolor{red}{arbitrarily} operating in different directions across islands, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Lailvaux2006, Nicholson2007}. Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances. Either way, new data would be needed to test these hypotheses. \textcolor{olive}{Visual modeling, for instance, would be a valuable follow-up analysis as it would indicate whether the differences we detected may be detectable by the organisms themselves (as has been done in \textit{Anolis} in \citealt{Leal2004, Fleishman2020}), but such approach requires irradiance profiles from the habitats in the field, which we presently do not have.}\\

>>>>

Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers)\sout{. Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (van de Schoot 2016 unpublished thesis), and (3) \textit{A. sagrei} is highly mobile (Kamath and Losos, 2018b), we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.} \textcolor{olive}{, and we would have expected gene flow to cause a more homogeneous distribution of color phenotypes within islands. While little is known about the cruising range of individuals from our study populations (but see \citealt{Steinberg2017, Kamath2018} for other systems), \textit{A. sagrei} are polygynandrous (both males and females mate with multiple mates, \citealt{Kamath2017, Kamath2018a, Kamath2018}), thus offering opportunity for gene flow, especially given that lizards were distributed continuously and at high densities within the islands we sampled. Consistent with that, while populations from different islands were monophyletic, individuals within islands were not monophyletic with respect to habitat based on mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis).}\\

@@

\sout{We found no evidence for a role of genetic drift in explaining the observed patterns either. First, \textit{A. sagrei} was distributed across the islands continuously, usually at relatively high population densities, rather than in small and isolated populations where drift might be expected to have a strong effect. Second, we found no evidence of isolation-by-distance except on Abaco, which was sampled at the largest geographical scale, with sites nearly a hundred kilometers apart from each other. Hence, while isolation-by-distance may explain long-range differences on this island, most of the differences among habitats across the rest of the sampling region are unlikely to be explained by genetic drift, as habitats were often in close proximity (less than 10km).}\\

@@

In this study, we found larger differences among than within islands, a pattern already reported and linked to climatic conditions \citep{Driessens2017} and to densities of predators and of anole congeners \citep{Vanhooydonck2009, Baeckens2018}. Differences among habitats within islands, however, are still difficult to account for. Remaining hypotheses may include, for example, runaway sexual selection \textcolor{olive}{(i.e. arbitrary preferences of females for some colors over others, \citealt{Andersson1994})} \sout{arbitrarily} operating in different directions across islands, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Lailvaux2006, Nicholson2007}. Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances. Either way, new data would be needed to test these hypotheses. \textcolor{olive}{Visual modeling, for instance, would be a valuable follow-up analysis as it would indicate whether the differences we detected may be detectable by the organisms themselves (as has been done in \textit{Anolis} in \citealt{Leal2004, Fleishman2020}), but such approach requires irradiance profiles from the habitats in the field, which we presently do not have.}\\
